Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-01-17 SCCAbstract_SDSS2022-knitr.Rnw
Fix bibliography and clarify some things.

\section{Current Progress and Future Work}
Our current design work will be broken down into four categories that have been identified in the following:
\begin{itemize}
\item User Flexibility 
\item Visual, Analytic, and Data Literacy
\item Data Design
\item Social Impact
\end{itemize}

In Figure 1, our dashboard will allow the user to select a town name, which will populate the  information in the maps related to necessary services, which include directions and distance to fire department, schools, post offices and hospitals. A table will be populated that will include information related to five other towns that have similar attributes, but the aspect of QoL will be included in the differences that help the town make a decision for the town.

@@

\subsection{Current Progress}
{\bf Statistical Techniques} Our research has incorporated statistical clustering methods, such as K-means and Hierarchical Clustering, to determine the similarities in the towns based on distances and services available. As a result, of our research partners we were actually able to use clustering methods to determine and identify missing data. We will continue to use clustering methods to identify the towns that will be best compare when the analyst interacts with our dashboard.

{\bf Data Science Techniques} Our team has been able to utilize direct API calls and web scraping tools to collect data necessary.

{\bf Dashboard Design Techniques} Our dashboard incorporates the following important dashboard techniques:
{\it User Flexibility} The aspects of user flexibility are directly based on the detail data adjustments; Adaptability: User and Display; Comparison support and Storytelling. For example, we want our users to use the data collected on a mass scale that have and should be summarized in detail but without losing anything.

{\it Visual, Analytic, and Data Literacy} The aspects in our dashboard development primary presents in our ability to ease of usage to the user with visual, data  and analytic literacy. We have little information on the users' adaptability to data-driven decisions.

{\it Data Design}. In our case, the challenges that arise in our dashboard will include data quality, data grouping, too much data, data sources, inaccurate data representation, and key performance indicators and metrics. For example, we have information from {\it Iowa Data} will have missing data can be pointed out by our users' that has not been captured by our data sources, this can be a recent closing of a school due to the pandemic.

{\it Social Impact} This aspect will be about making sure that the users are flexible to the following: data-driven thinking, technology resistance, dashboard aversion, and data sharing, security, and privacy. For example, we will encounter town users that will have the urge to compare their town to all types of towns that aren't similar. In our research, we will use statistical clustering methods and cater dashboard to show only towns that have similar components.

\subsection{Future Work}
Our dashboard design has created challenges that promote

We have five towns that have agreed to partner with the project to help with identify the best practices that are useful in the Iowa small and rural towns. As a result to our dashboard development, we will develop a feedback loop from the analyst in these five towns that will help provide our team with useful notes that will be adaptive to all of the small towns in Iowa. As a contrast, we will researchers with an extensive understanding of the small towns to be used as a counter example of dashboard adaptability. Our research, should be a

So dashboard challenges,
adaptation techniques, and user strategies are interconnected as
Figure 1 shows. The rationale is that if we are able to establish
a relationship between strategies to problems and we can detect
the strategies in real-time, we can adapt the user interface in realtime or in the next iteration. Then we can say that the adaption
techniques have a positive effect on the challenges.
\section{Acknowledgements}
We would like to thank our sponsors and grant provider NSF, LSAMP Inspire Program and Iowa League of Cities. We would like to thank our partners and team members at Iowa State University and University of Nebraska - Lincoln. We would like to thank the State of Iowa for providing Data along with our partners in the Rural Iowa communities.

>>>>

\section{Current Progress}
\svp{We leverage} the framework of Guided Discovery Learning (GDL) to guide \svp{the town leaders} to \svp{make discoveries using our interactive visualization}. \svp{This framework leverages hints, feedback, and other helpful information to guide users in interactive exploration}\cite{dedonno}.
% Stephen Few noted in {\it Information Dashboard Design: The effective visual communication of data} stated that visual orientation of dashboards is essential due to the speed at which perception \cite{few}.

%\begin{quotation}
%\smaller{The visual orientation of dashboards is important due to the speed of perception that’s usually required to monitor information. The faster you must assess what’s going on, the more you should rely on graphical means to display the information. \cite{few} }
%\end{quotation}

%Continue to cite this in the documentation but remove the quotation.

% A bit of background on Guided Discovery, which is defined as an environment where "students" are actively discovering knowledge, DeDonno states that we use GDL in our daily lives to make decisions by hints, feedback, and helpful information .
%\begin{quotation}
%\small In our daily lives, we may use a guided discovery process where learning is aided by hints, direction, feedback and other helpful information.  \cite{mayer} \cite{dedonno}
%\end{quotation}
% Few outlined common mistakes that are made in {\it Information Dashboard Design}, we have focused on making sure that the following three common mistakes on workplace dashboard development are avoided to best assist the town analyst to the best path of success with our solutions to those common mistakes:
% \begin{itemize}
% \item{ {\bf Choosing inappropriate display media} - Our work has minimized the number of tables and charts that could come from huge open-source data to incorporate highlighted key metrics and visualize data of five towns in parallel coordinate plots.}
% \item{ {\bf Ineffectively highlighting what’s important} - We wanted the town to be the center of focus for the user; as such, we added their information in the focal point in the center of the page. The comparison towns are only in the parallel coordinate plots to remind users that their town is the most important item in the dashboard.}
% \item{ {\bf Misusing or overusing color} - Our dashboard uses color to highlight the town in the parallel coordinate plot along with the leaflet maps that will be used in the theme of the Rural Shrink Smart Project}
% \end{itemize}

\svp{We provide users with a town-centric approach: their town is at the center of our application, and comparisons to other, similar towns are secondary. As it can be extremely difficult to predict which towns are optimal for comparison purposes (similar may involve population, region, economic indicators, sports rivalries, and any number of other variables), we allow users to modify a set of suggested comparison towns to indicate other towns of interest.}

\svp{As part of this project, we have assembled a large set of publicly available data (primarily from \url{data.iowa.gov}), but we are careful with how we present this data because it would be easy to overwhelm the user with small details that mask the bigger picture. We select a small subset of towns (out of the 999 towns in Iowa) and a small subset of variables of interest to start with, and then allow the user to increase the complexity of the display in accordance with their interest. This avoids some of the pitfalls of dashboard design that can easily lead to user overload \citep{few}.}

% The outlined list has been used as the guiding light to avoid pitfalls in developing dashboards in the workplace. With the amount of data, our dashboard could fall into the trap of over ineffectively highlighting what's important and choosing inappropriate display media and misusing color to ensure that the analyst's attention and all of the information is given to the user. This has been a significant hurdle when displaying the amount of data collected from open-source data for the small towns of Iowa.

Our current work incorporates the ideas of GDL, in which dashboard design has been used to display a panel with vital statistics, ensuring that the town's analyst small town is the center of focus. The vital statistics highlight the town's financial health and the shortest distances from the city center to critical resources, such as the fire department, hospital, post offices, and public schools. The dashboard includes Quality of Life (QoL) Metrics and the financial health of the small town. Based on the number of variables used to ensure that the dashboard is extensive and valuable, we implemented parallel coordinate plots to show comparisons between the focus small town with the five similar towns. A parallel (coordinates) plot allowed our town analyst to compare the essential features of several individual towns on a set of numeric and categorical variables collected. Historically, parallel coordinate plots have been used in air traffic control to help with collision avoidance algorithms but have been adapted in the statistical data visualization space, emphasizing three critical considerations: the order, the rotation, and the scaling of the axes. Our coordinate plots will display variables categorically and numerically by scaling the data.

We have implemented unsupervised statistical methods, such as K-means, Principal Component Analysis (PCA), and Hierarchical Clustering, to determine the five similar towns, based on distances and services available. We continue to use clustering methods to identify the towns that will be best compared when the analyst interacts with our dashboard. Using GDL, our dashboard makes sure that the focus small town is compared to other towns to find ways to improve their town using data-driven decisions. Our research partners have worked with five towns directly, who have agreed to partner with the project to help identify the best practices that are useful in the small and rural Iowa towns. As a result, we found out that the small towns do not want to be told what to do but would like to make decisions for those who have been a part of the community for generations. Our dashboard will hopefully make sure that the town's analysts are guided in the direction that will help the town's population shrink smart.

Our dashboard development has become more focused on the GDL such that a continual feedback loop from the town analyst in these five towns has helped provide our team with valuable notes that will be adaptive to all of the small towns in Iowa. Our team has captured valuable metrics on using the product allowing the quantitative and qualitative data collection to better user experience. This can promote challenges in towns feeling "studied" by people in cities outside of their own.

In Figure 1, our dashboard allows the user to select a town name, which will populate the information in the maps related to necessary services, including directions and distance to the fire department, schools, post offices, and hospitals. A table populates in the vital statistics sections that have information about the town's QoL Metrics and financial metrics, followed by a parallel coordinate plot that allows the town to see five towns that are similarly based on the most common variables in the towns, such as the distance of fire department.

@@

%To-Do: Write this in the way that we have done this because by the time that we present this will be done. 
%\subsection{Future Work}

\section{Acknowledgements}
We would like to thank our sponsors and grant provider NSF, LSAMP Inspire Program and Iowa League of Cities. We would like to thank our partners and team members at Iowa State University and University of Nebraska - Lincoln. We would like to thank the State of Iowa for providing Data along with our partners in the Rural Iowa communities.
